Towards Verifying the Bitcoin-S Library

### 2019-06-24 paper/paper.tex
update email
more intro

@@

\lstdefinestyle{scala}{
  language=scala,
%  basicstyle=\footnotesize\ttfamily,
  basicstyle=\scriptsize\ttfamily, % kai: temporary 
  breaklines=true,
  keywordstyle=\color{blue},
  commentstyle=\color{dkgreen},
  numbers=left,
  %frame=single, % Border around box
  %numbersep=-7pt,
  numberstyle=\color{gray},
  stringstyle=\color{mauve}
}

@@

\authorrunning{R. Boss et al.}
\institute{Bern University of Applied Sciences, CH-2501 Biel, Switzerland
\email{\{ramon.boss,kai.bruennler,anna.doukmak\}@bfh.ch}}

@@

The following example from the Stainless documentation
\cite{Stainless:documentation} demonstrates this. Notice how a
precondition is specified using the function \emph{require} and a
postcondition using \emph{ensuring}.

\begin{lstlisting}[style=scala]
  def factorial(n: Int): Int = {
      require(n >= 0)
      if (n == 0) {
        1
      } else {
        n * factorial(n - 1)
      }
  } ensuring(res => res >= 0)
\end{lstlisting}

Our program happens not to satisfy the specification. An overflow in
the 32-bit Int leads to a negative result for the input 17, as
Stainless reports in Figure~\ref{fig:failed}. Changing the type from
Int to BigInt will result in a successful verification.

@@

\paragraph{The Pure Scala Fragment.} Stainless supports only a subset
of Scala. They call this Pure Scala.  You can find the specification of Pure
Scala on the Stainless documentation website
\cite{Stainless:documentation} in the section
\href{https://epfl-lara.github.io/stainless/purescala.html}{Pure
  Scala}.
\todo{Kai: check the text

It supports Algebraic Data Type definitions in the form of 
Abstract Class and Case Class/Case Object,
inheritance of an Abstract Class by a Case Class/Case Object,
Object for grouping classes and functions,  
short-circuit interpretation of boolean expressions, 
Generics with invariant type parameters and simple hierarchy 
(subclasses define the same type parameters in the same order),
default values of function parameters,
Pattern matching, 
Values,
anonymous classes and more.

As an extension to Pure Scala Stainless supports some imperative features.
It supports 
declaring a variable in a local scope of a function, 
declaring a variable in trait without assigning a default value,
while loops,
case classes with \textit{var} arguments.

In particular Stainless does not support
Class definition,
inheritance by Object,
abstract type members,
variance annotations, upper and lower type bounds in generics,
variables in classes and objects,
private inner classes,
messages as a second argument in the Require function and more.}

In addition, Stainless has its own library with annotations,
reimplementation of some core data types, collections and input-output
functions and more. \todo{Kai: check the text.

This allows Stainless to map correctly operations 
to mathematical function inside of SMT solver.
Some elements of the library} are described on the website in the section
\href{https://epfl-lara.github.io/stainless/library.html}{Stainless
  Library}.  There are more details about the library in the source
code of Stainless on GitHub \cite{Stainless:github} in the folder
\href{https://github.com/epfl-lara/stainless/tree/master/frontends/library/stainless}{\textit{frontends/library/stainless}}.

@@

Some code in this section is copied or adapted from the Bitcoin-S-Core
transaction builder example \cite{BitcoinSCore:txbuilderexample}.
Bitcoin-S-Core has a bitcoin transaction builder class with the
following constructor:
\begin{lstlisting}[style=scala, basicstyle=\scriptsize\tt]
  BitcoinTxBuilder(
    destinations: Seq[TransactionOutput], // where to send the money
    utxos: BitcoinTxBuilder.UTXOMap,      // unspent transaction outputs
    feeRate: FeeUnit,                     // fee rate per byte
    changeSPK: ScriptPubKey,              // where to send the change
    network: BitcoinNetwork               // bitcoin network information
  ): Future[BitcoinTxBuilder]
\end{lstlisting}

>>>>

%\usepackage{newtxtext}
%\usepackage{newtxmath}
\usepackage[zerostyle=b,scaled=.9]{newtxtt}

@@

\lstdefinestyle{scala}{
  language=scala,
  basicstyle=\footnotesize\ttfamily,
%  basicstyle=\scriptsize\ttfamily, % kai: temporary 
  breaklines=true,
  keywordstyle=\color{blue},
  commentstyle=\color{dkgreen},
  numbers=left,
  %frame=single, % Border around box
  %numbersep=-7pt,
  numberstyle=\color{gray},
  stringstyle=\color{mauve}
}

@@

\authorrunning{R. Boss et al.}
\institute{Bern University of Applied Sciences, CH-2501 Biel, Switzerland
\email{ramon.boss@outlook.com, kai.bruennler@bfh.ch,\\anna.doukmak@gmail.com}}

@@

The example in Figure~\ref{fig:factorial} from the Stainless documentation
\cite{Stainless:documentation} demonstrates this. Notice how a
precondition is specified using the function \emph{require} and a
postcondition using \emph{ensuring}.

\begin{figure}
\begin{lstlisting}[style=scala]
  def factorial(n: Int): Int = {
      require(n >= 0)
      if (n == 0) {
        1
      } else {
        n * factorial(n - 1)
      }
  } ensuring(res => res >= 0)
\end{lstlisting}
	\caption{Factorial program with specification}
	\label{fig:factorial}
\end{figure}
Our program happens not to satisfy the specification. An overflow in
the 32-bit Int leads to a negative result for the input 17, as
Stainless reports in Figure~\ref{fig:failed}. Changing the type from
Int to BigInt will result in a successful verification.

@@

\paragraph{The Pure Scala Fragment.} The Scala fragment supported by
Stainless is described in the Stainless documentation
\cite{Stainless:documentation} in the section
\href{https://epfl-lara.github.io/stainless/purescala.html}{Pure
  Scala}.

It comprises algebraic data types in the form of abstract classes,
case classes and case objects, objects for grouping classes and
functions, boolean expressions with short-circuit interpretation,
generics with invariant type parameters, default values of function
parameters, pattern matching, local and anonymous classes and more.

In addition to Pure Scala Stainless also supports some imperative
features, such as using a (mutable) variable in a local scope of a
function and while loops. They turn out not to be relevant for the
current work.

What will turn out to be more relevant are the following Scala
features which Stainless does not support, such as: (concrete) class
definitions, inheritance by objects, abstract type members, variance
annotations, upper and lower type bounds in generics and private inner
classes.

In addition, Stainless has its own library of some core data types and
functions which need to be mapped correctly to functions inside of the
SMT solver that Stainless ultimately relies on. Those data types in
general do not have all the methods of the Scala data types. We will see an example of a method 
of the BigInt type that Scala provides but Stainless does not.
\todo{question viktor: where is BigInt??}

@@

Some code in this section is copied or adapted from the Bitcoin-S-Core
transaction builder example \cite{BitcoinSCore:txbuilderexample}.
Bitcoin-S-Core has a bitcoin transaction builder class with the
following constructor:
\begin{lstlisting}[style=scala]
  BitcoinTxBuilder(
    destinations: Seq[TransactionOutput], 
    utxos: BitcoinTxBuilder.UTXOMap,      
    feeRate: FeeUnit,                     
    changeSPK: ScriptPubKey,              
    network: BitcoinNetwork               
  ): Future[BitcoinTxBuilder]
\end{lstlisting}
